%{\LARGE Acknowledgements}\\
\section*{Acknowledgements}
During this research I received a large amount of support from a variety
of people. For this I am very grateful and I take this change to thank all of
them.\\

I'd like to thank my first supervisors Isaac Esteban for developing an 
idea that forms the start of this thesis.  Id like to thank him for sharing all his
knowledge about the field of multiple view geometry.  Although there was a large
difference in expectations on both sides, I learned a lot of him.  In my final
period he offered his time to gave me feedback about the parts which we did
together even though he was very busy writing his own thesis.\\

Furthermore I'd like to thank my second supervisor Frans Groen who gave me such professional supervision.  He
learned me the ingredients that a professional thesis needs.
Although he is almost retired, he came with very innovative ideas.  I loved the
sessions where we brainstormed about the techniques we used and where we
inspired each other by thinking out of the box.  Although he was very busy finishing
his job at UVA and TNO, he gave me full support and spend large amounts of his
time for my supervision.\\

Next I'd like to thank dr Piet Rodenburg who came from another research group
and was willing to join the committee on such a short notice.  Not only was the
subject new for him, also the time to
read this extensive thesis was very short.  dr Piet Rodenburg even took some
time of is holiday to read through a draft version of my thesis.\\

Furthermore I'd like to thank dr Arnoud Visser, who immediately said yes when I
invited him to the committee.  To participate my defense, Arnoud even arranged his
family to pick up his son from his last day of school.  Also thanks for
providing me latex formats and for reading my thesis in such a short time.  \\

I would like to thank my girlfriend Anne de Graaf who supported me during the
entire process.  I learned a lot from her motivation, her passion, her
gratefulness and her mentality to work hard without forgetting to enjoy life.
She used her excellent English skills and invested a large amount of her
expensive time to provide me spelling and grammar feedback.  Anne, thanks for
believing in me and for all the support that you gave me with so much love.\\

I would also like to thank my parents for their financial support, their latent
pressure, there advise and, most important, their love.\\

\newpage
A large part of this thesis was written in company with Toine van Asten who is my
best friend. He is one of the most motivated students I have ever met. Although
he studies at Delft, we worked large amount of time together. During his
exam periods and my thesis deadlines he learned me to take over hours as a
dessert.  Toine, thanks for supporting me with your commitment, insights, out of
the box thinking, positive energy, jokes, and your friendship.\\

I'd like to thank Bram Stoeller, my fellow classmate and, more important, warm
friend, for offering a large amount of his own time to read my thesis.\\

I'd like to thank Stefanie Kooistra who took a large amount of her time to
evaluate my thesis without expecting something in return.  Not only did she
highlight my spelling and grammar errors, she also learned me how to use the
grammar rules properly which was very helpful.\\

Furthermore, I'd like to thank Gineke Sietsma who trained me in Personal Leadership 
where I learned to develop my full potential.
