\section{Comments on this thesis}
%TODO
Dear Frans,

I write this in English so Isaac can read it also.
This in an update of my thesis chapter. Isaac read it and gave me feedback. His feedback and all feedback you gave me are incorporated except for:\\
-the test on different datasets\\
(the dataset is not ready yet)\\
-references\\
(I'll do this on the end)\\
-parametervalues Houghline\\
(I'll do this on the end)\\
-large introduction\\
Explanation:I expanded the introduction a bit more then my previous version but I think you
want more background information. I do this in a few chapters earlier which is
not included in this document. Therefore some backreferences to sections also
contain a questionmark.\\
\\
Furthermore, I processed many small changes but a few big changes are:\\
-I explicitly defined all assumptions.\\
-I wrote a new subsection about alternative roof-types.\\
-I updated the way line segments are associated with building parts and illustrated it with Figure 10.\\
\\
Because this new method of associating lines with walls is so much better and
more intuitive then my old method (see section 2), I am considering to
discard the old method.\\ 
However, I could also keep the old method and use it as
a test against the new one. What is your advice about this?\\

Would you give me feedback about the level of detail, where do I go in to much and where is more detail desired?\\
Isaac told me to make the future work section smaller, do you agree with this and if yes would you tell me which part you think we should drop?\\
Furthermore my discussion section is very short, I think it is best to add more discussion when the other datasets and results are ready. What do you think?\\
I have tried to write down some mathematics but didn't do it everywhere. Could you give me feedback where to put (more) formulas or symbols and where not?\\
I think I need to put my work in context with the work of others. But on this specific part I can't find any work. How do I deal with that?\\
Do I remember it right that you are leaving for 1 month beginning at half september?\\
Would be nice if we can discuss your feedback before then, but that will depend on your schedule.\\
Please let me know when the feedback is done and lets make an appointment. I think Isaacs presence will not be necesarry, right Isaac?\\
I am available most of the days. Please let me know when you are available.\\

Thanks in advance for reading my thesis chapter and giving me useful feedback.\\

Kind regards,\\
Tjerk Kostelijk\\
mailtjerk@gmail.com
